---
title: "Visualizations Assignment"
author: "Chase LaDue"
date: "4/3/2018"
output: html_document
---

First, I'll start with the code that was given in `working_example.Rmd`.

```{r package-data}

library(tidyverse) # loading package

sjer_veg <- read_csv("data/D17_2013_SJER_vegStr.csv") # loading data

```

Now, let's plot the graph that we'll be starting with.

```{r base-graph}

g <- sjer_veg %>% 
  select(maxcanopydiam, stemheight, livingcanopy) %>% 
  ggplot(aes(x = stemheight, y = maxcanopydiam)) + 
    geom_point(aes(color = livingcanopy)) +
    scale_color_gradient(low = "red", high = "green")

g # view graph

```

Let's start off by providing appropriate labels for the title, axes, and legend.

```{r labels}

# data description is available in file "veg_metadata_desc_DataInstitute2016.csv"

# Future self: to get rid of error message of replacing color gradient argument, let's eliminate that from g real quick:

g <- sjer_veg %>% 
  select(maxcanopydiam, stemheight, livingcanopy) %>% 
  ggplot(aes(x = stemheight, y = maxcanopydiam)) + 
    geom_point(aes(color = livingcanopy)) #+
#    scale_color_gradient(low = "red", high = "green")

# That's better. Now we can proceed.

g.1 <- g +
  scale_color_gradient(low = "red", high = "green", name = "% of canopy alive") + # legend label
  xlab("Distance from ground to highest stem (m)") + # added x-axis label
  ylab("Maximum canopy diameter (m)") + # added y-axis label
  ggtitle("Relationship between\nCanopy Diameter and Distance to Highest Stem") # graph title

g.1 # view updated graph
```

Now readers can better understand what's being plotted and compared. Let's change the look of this graph now by changing the theme and cleaning up unnecessary elements.

```{r theme-change}

custom.theme <- theme(panel.grid.major = element_blank(), # these lines gets rid of distracting grid in background
                      panel.grid.minor = element_blank(),
                      panel.background = element_rect(fill="white"),
                      axis.line = element_line(colour="black"), # this changes axes lines to black
                      axis.text.x = element_text(colour="black", size=10), # changes axes text (numbers)
                      axis.text.y = element_text(colour="black", size=10),
                      axis.title.x = element_text(colour="black", size=12, face="bold"), # changes axes labels
                      axis.title.y = element_text(colour="black", size=12, face="bold"),
                      plot.title = element_text(hjust=0.5), # centers plot title
                      legend.title = element_text(face="bold", size=10), # changes legend title text
                      legend.title.align = 0.5, # centers legend title
                      legend.text = element_text(size=8), # changes legend text
                      legend.background = element_rect(fill="#F6F6F6", colour = "black")) # adds background and border to legend

g.2 <- g.1 + custom.theme

g.2
```

With cleaned-up graph, it is difficult to estimate the values of the points on the graph. A solution to this problem is to increase the frequency of tick marks.

```{r tick-marks}

g.3 <- g.2 + 
  scale_y_continuous(breaks=seq(0,35,by=5)) + # sets tick marks every 5 units on y axis
  scale_x_continuous(breaks=seq(0,30,by=2)) # sets tick marks every 2 units on x axis

g.3

```

